\subsection{XML Streaming}\label{sec:xml} % Sherif


In 2002, Diao et al.~\cite{diao_et_al_2002,diao2003high} presented YFilter as a system that implements
continuous queries over XML streaming data using the XPath language~\cite{clark_derose_1999}. YFilter applies
a multi-query optimization mechanism that uses a single
finite state machine to represent and evaluate several XPath
expressions. In particular, YFilter exploits the commonality among path queries by merging the common prefixes
of the paths so that they are processed at most once. This  shared processing mechanism provides significant performance
improvement by avoiding redundant processing for duplicate path expressions. In order to handle value-based predicates that address contents of elements, YFilter applies two alternative approaches. The first approach approach evaluates the  predicates once the addressed elements are read from a document, while the other approach postpones predicate evaluation until the corresponding path expression has been entirely matched.
XSQ~\cite{peng_chawathe_2003} is another system that uses XPath for querying streaming XML data. It is based on a hierarchical arrangement of pushdown transducers
augmented with buffers.

\iffalse
\begin{figure}[!h]
\begin{lstlisting}
CREATE ~\textit{cq\_name}~
         ~\textit{xml\_ql\_query}~
     DO ~\textit{action}~
  {START ~\textit{start\_time}~}
  {EVERY ~\textit{time\_interval}~}
  {EXPIRE ~\textit{expiration\_time}~}
\end{lstlisting}
\vspace*{-4mm}
\caption{\label{fig:Niagra}NiagaraCQ code example.}
\end{figure}
\fi

NiagaraCQ~\cite{chen_et_al_2000} is a 
continuous query engine for XML data that supports incremental evaluation
and considers only the changed portion of each updated XML file. In particular, it
supports two types of continuous queries that differ in the criteria
triggering their execution: \emph{change-based} queries, which trigger
as soon as new relevant data becomes available, and \emph{timer-based}
queries, which trigger only at specified time intervals.  The syntax of NiagaraCQ is
based on the XML-QL language~\cite{deutsch1999query}.  \iffalse It provides a command
language for creating continuous queries that follows the form
illustrated in Figure~\ref{fig:Niagra}. Using this command language,
users can implement continuous queries that combines an ordinary
XML-QL query with additional time information.  The
\textsf{\small\textit{xml\_ql\_query}} becomes effective at the
defined \textsf{\small\textit{start\_time}}.  The
\textsf{\small\textit{time\_interval}} indicates how often the query
will be executed. A query is timer-based if its
\textsf{\small\textit{time\_interval}} is non-zero; otherwise, it is
change-based.  The continuous query will be deactivated after its
\textsf{\small\textit{expiration\_time}}. The
\textsf{\small\textit{action}} triggers once the results of the XML-QL
query expression is returned.  A key optimization mechanism used in
the NiagaraCQ system is grouping the execution of similar queries to
minimize redundant work.\fi


Mendell et al.~\cite{mendell_et_al_2012} presented an approach for extending a general-purpose streaming language, SPL~\cite{hirzel_schneider_gedik_2017}, 
with a declarative syntax for XML processing.
In particular,  they introduced the  \emph{XMLParse} operator which allows the user 
to not only filter the XML document but also specify a transformation of the returning XML data 
 to amenable format that can be used within the general-purpose streaming system.

